---
output:
  word_document: default
  html_document: default
---
# **Graduate Rotational Internship Program (The Spark Foundation)**
# TASK 2 (To Explore Supervised Machine Learning)

First let us start with importing required libraries and data.
PS: please set your current directory to the location where your data-set is, by using setwd("").

```{r}
#Loading libraries
library(tidyverse)
library(caret)
library(ggplot2)

#reading data from the URL
data <- read.csv(url("https://raw.githubusercontent.com/AdiPersonalWorks/Random/master/student_scores%20-%20student_scores.csv"))
```

We have to predict the percentage of marks that a student is expected to score based upon the number of hours they studied.

Let's first observe our data.

```{r}
#getting the summary of the data
summary(data)
```

So there are no *NA* values.

Now, let's visualize our data.

```{r}
#plotting the data
data %>% ggplot(., aes(Hours, Scores)) + geom_point(size = 3) + scale_x_continuous(breaks = c(0:10)) + 
  scale_y_continuous(n.breaks = 10) +  theme_classic()
```

It is clearly visible that both variables are positively correlated (cor(Hours, Scores) = `r cor(data$Hours, data$Scores)`). We can easily fit a linear model to the data.

Before fitting the regression let's split the data into training and test set.

```{r}
#splitting the data into the training and test set
set.seed(122, sample.kind = "Rounding")
test_index <- createDataPartition(data$Scores, times = 1, p = 0.3, list = FALSE)
train_set <- data[-test_index, ]
test_set <- data[test_index, ]
```

Now, fitting linear regression to the data, and predicting the model on the test set.

```{r}
#fitting linear reg.
fit <- lm(data = train_set, formula = Scores ~.)
summary(fit)
pred <- predict(fit, test_set)
```

We are getting the goodness of fit as 0.9539 or we can say that our model has explained 95.39% of the variation in the percentage of marks scored is explained by the no. hours a student is studying, with a root mean square error of `r RMSE(pred, test_set$Scores)`

Visualizig our fit.

``` {r}
plot(data, main = "Regression Plot")
abline(fit, col = "dark red")
```

Determining the percentage of marks obtained by studying for 9.25 hours.

```{r}
p <- data.frame(Hours = 9.25, Scores = "")
predict(fit, p)
```

So the predicted score is `r predict(fit, p)`.